---
title: "Data Preparation"
author: "James Watson"
date: "4/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = T, cache.comments = FALSE,
                      echo = TRUE, include = TRUE, 
                      fig.width = 8, fig.height = 8,
                      fig.pos = 'H',dev = 'png', dpi = 300)
library(missForest)
library(dplyr)

# Variables that control Markdown run
RE_RUN_IMPUTATION = F
```



## Overview

This script puts together clinical data from multiple studies in order to estimate probabilities that Kenyan children enrolled in a large observational cohort had severe malaria.

This script can only be run with access to original datasets - need to contact XXXXX.

Data used to build the reference model of severe malaria:

* AQ Vietnam (platelets and white counts): adults
* AAV Vietnam (platelets and white counts): adults
* MORU database (white counts and platelets): kids
* FEAST (white counts, platelets, Hrp2): kids


## Kenyan dataset

Load data
```{r}
load('../RData/kemri_case_data.RData')
load('../RData/kemri_control_data.RData')

kemri_case_data$sickle_trait = as.numeric(kemri_case_data$sickle==2)
kemri_control_data$sickle_trait = as.numeric(kemri_control_data$sickle==2)
writeLines('Distribution of the number of WT alleles (A) at rs_334 in the severe malaria cases:')
table(kemri_case_data$sickle-1, useNA = 'ifany')
writeLines('Distribution of the number of WT alleles (A) at rs_334 in the population controls:')
table(kemri_control_data$sickle-1, useNA = 'ifany')
```


Do imputation for missing variables
```{r Missing_data_KEMRI}
K_imputes = 10

dat = kemri_case_data

dat$baseexc[dat$baseexc > 30]= NA
dat$BD = -dat$baseexc

dat$log10_platelets = log10(dat$platelet)
dat$log10_wbc = log10(dat$wbc)
dat$log10_creatinine = log10(dat$creat)
dat$log10_creatinine[is.infinite(dat$log10_creatinine)] = NA

dat = dat[, -which(colnames(dat)=='baseexc')]
vars_interest = colnames(dat)
writeLines('All available variables:')
print(colnames(dat))
vars_interest = vars_interest[!vars_interest%in% 
                                c('ageyr','IID','case_control','thal',
                                  'bacteria',"sample_code",
                                  'sickle','g6pd_202','g6pd_376',
                                  "pat_id","source_code","parasite_gn",
                                  'bcs_total', 'bcs_eyes','bcs_motor',
                                  'bcs_verbal', 'platelet', 'wbc','creat')]

Missing_data = is.na(dat[,vars_interest])
par(las=1, mar=c(4,9,2,2))
image(x = 1:nrow(dat), y = 1:length(vars_interest), 
      Missing_data, yaxt='n',ylab='',xaxt='n', xlab='')
axis(2, at = 1:length(vars_interest), 
     labels = tolower(vars_interest),cex=.7)

writeLines(sprintf('There is clinical data on %s individuals', nrow(dat)))
apply(kemri_case_data, 2, function(x) mean(is.na(x)))

# redo this with platelets and wbc on log scale
if(RE_RUN_IMPUTATION){
  set.seed(8758768)
  registerDoParallel(cores = 6)
  
  imputed_data_KEMRI = list()
  my_dat = dat[,vars_interest]
  discrete_vars = c('sex','ethnic','died','prostration','cerebral_malaria',
                    'respiratory_distress','bloodculture_pos','severe_malaria_anaemia')
  for(cc in discrete_vars){
    my_dat[,cc] = as.factor(my_dat[,cc])
  }
  for(k in 1:K_imputes){
    out_impute = missForest(xmis = my_dat, variablewise = T, 
                            ntree = 200,
                            decreasing = T,verbose = T, 
                            parallelize = 'forests')
    out_impute$ximp$hypo_glycaemic = out_impute$ximp$glucose<2.2
    imputed_data_KEMRI[[k]] = out_impute$ximp
  }
  save(imputed_data_KEMRI, file = '../RData/Imputed_Datasets_KEMRI.RData')
} else {
  load('../RData/Imputed_Datasets_KEMRI.RData')
}
```



## Reference model datasets


AQ study - impute missing data
```{r Missing_data_AQVietnam}
# Load the AQ Vietnam data
aq_dat = readstata13::read.dta13('~/Dropbox/Datasets/AQ study/Vietnam AQ.DTA')
aav_dat = read.csv('~/Dropbox/Datasets/AllMalariaDataBackUp/AAV_Vietnam/AAV.csv')
aav_dat[aav_dat==999999] = NA
aq_dat_training = data.frame(platelets_log10 = log10(aq_dat$admplat/10^3),
                             wbc_log10 = log10(aq_dat$admwbc/10^3),
                             parasitaemia_log10 = log10(aq_dat$admpct),
                             hct = aq_dat$admhct,
                             age = aq_dat$age,
                             outcome = aq_dat$outcome,
                             gcs = aq_dat$admgcs,
                             creatinine_log10 = log10(aq_dat$admcrea),
                             sex = aq_dat$sex)
aav_dat_training = data.frame(platelets_log10 = log10(aav_dat$admplt/10^3),
                              wbc_log10 = log10(aav_dat$admwbc/10^3),
                              parasitaemia_log10 = log10(aav_dat$admparasitemia),
                              hct = aav_dat$initialhct,
                              age = aav_dat$age,
                              outcome = aav_dat$mortalityoutcome,
                              gcs = aav_dat$gcstotal,
                              creatinine_log10 = log10(aav_dat$admcrea),
                              sex = aav_dat$sex)
Vietnam_training = rbind(aq_dat_training, aav_dat_training)

apply(Vietnam_training, 2, function(x) round(100*mean(is.na(x)),1))
par(las=1, mar = c(2,9,2,2))
image(x = 1:nrow(Vietnam_training), 
      y = 1:ncol(Vietnam_training), is.na(Vietnam_training),
      yaxt='n',ylab='',xaxt='n', xlab='')
axis(2, at = 1:ncol(Vietnam_training), labels = tolower(colnames(Vietnam_training)),cex=.7)

if(RE_RUN_IMPUTATION){
  set.seed(8758768)
  registerDoParallel(cores = 6)
  
  imputed_data_Viet = list()
  my_dat = Vietnam_training
  discrete_vars = c('sex','outcome')
  for(cc in discrete_vars){
    my_dat[,cc] = as.factor(my_dat[,cc])
  }
  for(k in 1:K_imputes){
    out_impute = missForest(xmis = my_dat, variablewise = T, 
                            ntree = 200,
                            decreasing = T,verbose = T, 
                            parallelize = 'forests')
    imputed_data_Viet[[k]] = out_impute$ximp
  }
  save(imputed_data_Viet, file = '../RData/Imputed_Datasets_Vietnam.RData')
} else {
  load('../RData/Imputed_Datasets_Vietnam.RData')
}

# make an imputed dataset just for platelets and white counts
imputed_Vietam_minimal = list()
for(k in 1:K_imputes){
  imputed_Vietam_minimal[[k]] = 
    imputed_data_Viet[[k]][, c('platelets_log10','wbc_log10','age')]
}
save(imputed_Vietam_minimal, file = 'Inputs/imputed_Vietam_minimal.RData')
writeLines(sprintf('We include a total of %s adults from Vietnam',nrow(Vietnam_training)))
```


MORU core malaria (severe patients only)
```{r MORU core malaria}
# Core Malaria - MORU: children with severe malaria
# prepared using file extract_data.R in the hrp2 folder
load(file = '../RData/Core_malaria_hrp2_platelets.RData')
# filter out one super low platelet count (probably data entry mistake)
dat_core = core_data[core_data$platelets>1, ]
print(table(dat_core$country))
hist(dat_core$age, breaks = seq(0,80,by=5))
writeLines(sprintf('We include a total of %s patients from the Bangladesh and Thailand',nrow(dat_core)))
```


FEAST
```{r}
load(file = '../RData/feast_platelet_white_count_data.RData')
# select the patients with high HRP2 and parasite counts as training data for the reference model
ind_feast_training = !is.na(dat_feast$platelets) & 
  !is.na(dat_feast$white_count) & 
  !is.na(dat_feast$hrp2) & 
  !is.na(dat_feast$parasitaemia) & 
  dat_feast$hrp2 > 1000 & dat_feast$parasitaemia > 1000 
writeLines(sprintf('The diagnostic model training data includes a total of %s patients from the FEAST study',
                   sum(ind_feast_training)))

ind_platelet_hrp2 = !is.na(dat_feast$platelets) &
  !is.na(dat_feast$hrp2)
xx = dat_feast[ind_platelet_hrp2, c('platelets','hrp2','parasitaemia')]
write.csv(xx, file = 'Inputs/FEAST_platelets_hrp2.csv',
          row.names = F)
writeLines(sprintf('We use a total of %s patients from the FEAST study to show the relationship between PfHRP2 and platelet counts',
                   sum(ind_platelet_hrp2)))
```


We use the data from FEAST to estimate the age-related trend in white counts
```{r age_model}
# Age model of white counts
wbc_age_model = mgcv::gam(formula = log10_wbc ~ s(age), 
                          data = data.frame(age=dat_feast$age,
                                            log10_wbc=log10(dat_feast$white_count)))
summary(wbc_age_model)
# save model for reproducibility
save(wbc_age_model, file = 'Inputs/wbc_age_model')

par(las=1, cex.lab=1.5, cex.axis=1.5)
plot((0:120)/12, predict(wbc_age_model, newdata=data.frame(age=0:120)),
     type='l',lwd=3, xlab='Age (years)', ylab ='White blood cell count (x1000 per uL)',
     yaxt='n', ylim = range(log10(dat_feast$white_count),na.rm = T))
axis(2, at = seq(0,2, by=0.5), labels = round(10^seq(0,2, by=0.5),1))
points(jitter(dat_feast$age/12), log10(dat_feast$white_count), pch=20)


# estimate the expected baseline white count in severe disease in children older than 60 months
baseline_log10_wbc = mean(log10(dat_feast$white_count[dat_feast$age>60]), 
                          na.rm = T)
dat_feast$wbc_log10 = log10(dat_feast$white_count)

# calculate an age-corrected white count for FEAST children
dat_feast$wbc_log10_corrected = dat_feast$wbc_log10

ind = dat_feast$age<60 & !is.na(dat_feast$white_count)
dat_feast$wbc_log10_corrected[ind] = 
  dat_feast$wbc_log10_corrected[ind]-
  (predict(wbc_age_model, 
           newdata=data.frame(age=dat_feast$age[ind]))-baseline_log10_wbc)

dat_feast_training = 
  data.frame(platelets_log10 = log10(dat_feast$platelets)[which(ind_feast_training)],
             wbc_log10 = dat_feast$wbc_log10[which(ind_feast_training)],
             wbc_log10_corrected = dat_feast$wbc_log10_corrected[which(ind_feast_training)])

# calculate an age-corrected white count for Core Malaria children
ind = !is.na(dat_core$age)&dat_core$age<5
sum(ind)
dat_core$log10_wbc = log10(dat_core$wbc)
dat_core$platelets_log10 = log10(dat_core$platelets)
dat_core$log10_wbc_corrected = log10(dat_core$wbc)
dat_core$log10_wbc_corrected[ind] = dat_core$log10_wbc[ind]-
  (predict(wbc_age_model, newdata=data.frame(age=dat_core$age[ind]*12))-baseline_log10_wbc)

# calculate an age-corrected white count for Kenyan children
ind = !is.na(kemri_case_data$agemths) & kemri_case_data$agemths<60
# extract an imputed wbc set
kemri_case_data$log10_wbc_imputed = imputed_data_KEMRI[[1]]$log10_wbc
kemri_case_data$log10_platelet_imputed = imputed_data_KEMRI[[1]]$log10_platelets
kemri_case_data$log10_wbc_imputed_corrected = kemri_case_data$log10_wbc_imputed
kemri_case_data$log10_wbc_imputed_corrected[ind] = kemri_case_data$log10_wbc_imputed[ind]-
  (predict(wbc_age_model, newdata=data.frame(age=kemri_case_data$agemths[ind]))-baseline_log10_wbc)
```

## Compare platelet counts and white counts across datasets

QQ plots comparing studies
```{r qqplot_SM_studies_wbc}
par(mfrow=c(2,2), las= 1, family='serif', cex.lab=1.5, cex.axis=1.5)
qqplot(imputed_Vietam_minimal[[1]]$wbc_log10, # no crrection needed s all adults
       dat_feast_training$wbc_log10_corrected, 
       xlim=c(0,2), ylim=c(0,2),panel.first=grid(),
       xlab='White count (Vietnamese adults)',
       ylab='White count (FEAST)', xaxt='n',yaxt='n',pch=16)
axis(side = 1, at = 0:2,  labels = 10^(0:2))
axis(side = 1, at = log10(c(seq(1,10,by=1), seq(10,100,by=10))), 
     tick = T, labels = F)
axis(side = 2, at = 0:2,  labels = 10^(0:2))
axis(side = 2, at = log10(c(seq(1,10,by=1), seq(10,100,by=10))), 
     tick = T, labels = F)
lines(0:5, 0:5)


qqplot(dat_core$log10_wbc_corrected, 
       dat_feast_training$wbc_log10_corrected, 
       xlab='White count (Bangladesh/Thailand)',
       xlim=c(0,2), ylim=c(0,2),panel.first=grid(),
       ylab='White count (FEAST)', xaxt='n',yaxt='n',pch=16)
axis(side = 1, at = 0:2,  labels = 10^(0:2))
axis(side = 1, at = log10(c(seq(1,10,by=1), seq(10,100,by=10))), 
     tick = T, labels = F)
axis(side = 2, at = 0:2,  labels = 10^(0:2))
axis(side = 2, at = log10(c(seq(1,10,by=1), seq(10,100,by=10))), 
     tick = T, labels = F)
lines(0:5, 0:5)


qqplot(c(dat_core$log10_wbc_corrected,imputed_data_Viet[[1]]$wbc_log10), 
       dat_feast_training$wbc_log10_corrected, 
       xlab='White count (Asia)',
       xlim=c(0,2), ylim=c(0,2),panel.first=grid(),
       ylab='White count (FEAST)', xaxt='n',yaxt='n',pch=16)
axis(side = 1, at = 0:2,  labels = 10^(0:2))
axis(side = 1, at = log10(c(seq(1,10,by=1), seq(10,100,by=10))), 
     tick = T, labels = F)
axis(side = 2, at = 0:2,  labels = 10^(0:2))
axis(side = 2, at = log10(c(seq(1,10,by=1), seq(10,100,by=10))), 
     tick = T, labels = F)
lines(0:5, 0:5)
```


```{r qqplot_SM_studies_platelets}
#************* Platelets ***************
par(mfrow=c(2,2), las= 1, family='serif', cex.lab=1.5, cex.axis=1.5)

qqplot(imputed_Vietam_minimal[[1]]$platelets_log10, 
       dat_feast_training$platelets_log10, 
       xlab='Platelet count (Vietnamese adults)',
       panel.first=grid(),
       xlim = c(1,3), ylim=c(1,3),
       ylab='Platelet count (FEAST)', pch=16, xaxt='n',yaxt='n')
axis(side = 1, at = 1:3, labels = 10^(1:3))
axis(side = 1, at = log10(c(seq(10,100,by=10), seq(100,1000,by=100))), 
     tick = T, labels = F)
axis(side = 2, at = 1:3, labels = 10^(1:3))
axis(side = 2, at = log10(c(seq(10,100,by=10), seq(100,1000,by=100))), 
     tick = T, labels = F)
lines(0:5, 0:5)

qqplot(log10(dat_core$platelets), 
       dat_feast_training$platelets_log10, 
       panel.first=grid(),
       xlim = c(1,3), ylim=c(1,3),
       xlab='Platelet count (Bangladesh/Thailand)',
       ylab='Platelet count (FEAST)', pch=16, xaxt='n',yaxt='n')
axis(side = 1, at = 1:3, labels = 10^(1:3))
axis(side = 1, at = log10(c(seq(10,100,by=10), seq(100,1000,by=100))), 
     tick = T, labels = F)
axis(side = 2, at = 1:3, labels = 10^(1:3))
axis(side = 2, at = log10(c(seq(10,100,by=10), seq(100,1000,by=100))), 
     tick = T, labels = F)
lines(0:5, 0:5)


qqplot(c(log10(dat_core$platelets), imputed_Vietam_minimal[[1]]$platelets_log10),
       dat_feast_training$platelets_log10, 
       xlim = c(1,3), ylim=c(1,3),
       panel.first=grid(),
       xlab='Platelet count (Asia)',
       ylab='Platelet count (FEAST)', pch=16, xaxt='n',yaxt='n')
axis(side = 1, at = 1:3, labels = 10^(1:3))
axis(side = 1, at = log10(c(seq(10,100,by=10), seq(100,1000,by=100))), 
     tick = T, labels = F)
axis(side = 2, at = 1:3, labels = 10^(1:3))
axis(side = 2, at = log10(c(seq(10,100,by=10), seq(100,1000,by=100))), 
     tick = T, labels = F)
lines(0:5, 0:5)
```

QQ plots looking at marginal normality
```{r qq_normality, fig.height=6, fig.width=10}
k=1
dat_Training = rbind(imputed_Vietam_minimal[[k]][, c('platelets_log10', 'wbc_log10')],
                      rename(dat_feast_training[, c('platelets_log10', 'wbc_log10_corrected')],
                             wbc_log10=wbc_log10_corrected),
                      rename(dat_core[, c('platelets_log10', 'log10_wbc_corrected')],
                             wbc_log10=log10_wbc_corrected))
writeLines('Dimensions of training dataset:')
print(dim(dat_Training))

par(las=1, bty='n', mfrow=c(1,2), family='serif', 
    cex.lab=1.5, cex.axis=1.5)
qqnorm(scale(dat_Training$platelets_log10), pch=20, 
       panel.first=grid(), main = 'Log10 platelet counts')
lines(-10:10, -10:10, lwd=2, lty=2)
qqnorm(scale(dat_Training$wbc_log10), pch=20, 
       panel.first=grid(), main = 'Log10 white counts')
lines(-10:10, -10:10, lwd=2, lty=2)
```



Make curated dataset
```{r}
dat_kenya = kemri_case_data[ , c('agemths',
                                 'platelet', # observed platelets
                                 'wbc',      # observed white counts
                                 'log10_platelet_imputed', # imputed white counts (log10 scale)
                                 'log10_wbc_imputed', # imputed white counts (log10 scale)
                                 'log10_wbc_imputed_corrected')]  #with age correction
dat_kenya$HbAS = kemri_case_data$sickle_trait

dim(dat_kenya)
dim(dat_Training)
dim(dat_feast_training)
save(dat_kenya, dat_Training, dat_feast_training, dat_core, 
     file = 'Inputs/curated_modelling_dataset.RData')
```

